---
title: "An Analysis of the Gender Pay Gap in UK Univerisities"
author: "Matthew Knowles"
date: "Semester 1, 2022/23"
output: 
    html_document:
        toc: true
        toc_float: true
        number_sections: true
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/matthew/Documents/EDA_CS1")
library(ggplot2)
library(dplyr)
library(mice)

# Load Data
GPG1718 <- read.csv(file = "data/GPG1718.csv")
GPG1819 <- read.csv(file = "data/GPG1819.csv")
GPG1920 <- read.csv(file = "data/GPG1920.csv")
GPG2021 <- read.csv(file = "data/GPG2021.csv")
GPG2122 <- read.csv(file = "data/GPG2122.csv")
unis <- read.csv("data/universities.csv")

# Add a year coolumn to each data
GPG1718$year <- 2017
GPG1819$year <- 2018
GPG1920$year <- 2019
GPG2021$year <- 2020
GPG2122$year <- 2021

# Compile data
GPG <- rbind(GPG1718, GPG1819, GPG1920, GPG2021, GPG2122)

# Just University GPG data
uniGPG <- GPG %>%
    filter(EmployerName %in% unis$EmployerName)

# Get the count of reporting
uni_year_count <- uniGPG %>%
    count(EmployerName)

# Plot the density of reporting frequency
uni_year_count_density <- density(uni_year_count$n)
plot(uni_year_count_density)
```

# Introduction

Here, we will take a look at the Gender Pay Gap in the UK. Specifically, in UK universities.

# Data Overview

## Missing Data

Missing data is a common problem in data analysis. 

```{r, echo = FALSE, results = 'hide'}
md.pattern(uniGPG, rotate.names = TRUE)
```

We see that the two missing datapoints are the difference in mean and median bonus percent.

```{r, echo = FALSE}
uniGPG <- na.omit(uniGPG)
```

# Densities

We first look at how a few variables are distributed. First, the average hourly wage difference as a percentage. The way this is presented shows older data in a darker colour, with the magenta line being the most recent full year of data. This is done to highlight how the denisty has changed over time up to the most recent year.

```{r, echo = FALSE}
ggplot(uniGPG, aes(x = DiffMeanHourlyPercent)) +
    geom_density(data = filter(uniGPG, year == 2017), color = "#19234d", show.legend = TRUE) +
    geom_density(data = filter(uniGPG, year == 2018), color = "#2d3d86", show.legend = TRUE) + 
    geom_density(data = filter(uniGPG, year == 2019), color = "#4057bf", show.legend = TRUE) + 
    geom_density(data = filter(uniGPG, year == 2020), color = "#7989d2", show.legend = TRUE) + 
    geom_density(data = filter(uniGPG, year == 2021), color = "magenta", show.legend = TRUE)
```

What about bonuses?

```{r, echo = FALSE }
ggplot(uniGPG, aes(x = DiffMeanBonusPercent)) +
    geom_density(data = filter(uniGPG, year == 2017), color = "#19234d", show.legend = TRUE) +
    geom_density(data = filter(uniGPG, year == 2018), color = "#2d3d86", show.legend = TRUE) + 
    geom_density(data = filter(uniGPG, year == 2019), color = "#4057bf", show.legend = TRUE) + 
    geom_density(data = filter(uniGPG, year == 2020), color = "#7989d2", show.legend = TRUE) + 
    geom_density(data = filter(uniGPG, year == 2021), color = "magenta", show.legend = TRUE)
```
